\vspace{3cm} \Huge{\futhark}\\ \Large{Enhancing performance by rewriting kernels from Python/OpenCL to Futhark}

### 2017-11-08 report/draft.tex
dd

\documentclass[11pt]{article}
\usepackage[a4paper, hmargin={2.8cm, 2.8cm}, vmargin={2.5cm, 2.5cm}]{geometry}
\usepackage{eso-pic} % \AddToShipoutPicture
\usepackage{graphicx} % \includegraphics
\usepackage{listings} % \includegraphics
\usepackage{pgfplots}
\usepackage{hyperref}

@@

\title{
  \vspace{3cm}
  \Huge{Optimizing SasView models} \\
  \Large{Enhancing performance by rewriting kernels from OpenCL/OpenMP to Futhark}
}

\author{
  \Large{Mikkel Storgaard Knudsen}
  \\ \texttt{mikkelstorgaard@gmail.com} \\
  \\
  \Large{Supervisor}
  \\ Martin Elsman
  \\ \texttt{mael@di.ku.dk} \\
}

@@

\newcommand{\sasmodels}{\textit{SasModels}}
\newcommand{\sasview}{\textit{SasView}}
\newcommand{\iq}{\textit{I(q)}}
\newcommand{\foriq}{\texttt{for\_iq}}
\newcommand{\hasiq}{\texttt{has\_iq}}
\begin{document}

@@

\section{Introduction}
\subsection{Goal and motivation}
\textit{SasView} is a Small Angle Scattering (SAS) analysis package for the analysis of
1D and 2D scattering data directly in inverse space.
As \textit{SasView} is given data, it performs its analyses by running the input data
through theory models provided by the standalone module \textit{Sasmodels}.
\\
\\
The \textit{Sasmodels} models utilizes a set of kernels, which are implemented
in both OpenCL and OpenMP.
However, benchmark
comparisons\footnote{https://futhark-lang.org/performance.html} between the
parallel algorithm library \textit{Thrust} (implemented in C++/CUDA), and the
array programming language Futhark, shows that there might severe speed
improvements to gain by running these model calculations in Futhark, instead
of in OpenCL and OpenMP.
\\
\\
The goal of this project is to explore the potential speed gains obtained by
rewriting, and testing the resulting speed improvements of, a subset of the
\textit{Sasmodels} models.
The subset will be chosen so that there are at least
one ``simple'' model, one model of intermediate complexity, and one complex
model in it.
\\
\\
If rewriting the models results in a significant\footnote{I.e. a speed increase
  in order of at least a magnitude}, speedup, it will reasonable to believe that 
  the rest of the models in \textit{Sasmodels} can be rewritten for a similar
  speedup.

\section{Demonstration of use of finished project}
As \sasmodels is rarely used on its own, but rather as the backbone of 
\sasview's calculation backend, I am not going to write much about 
\sasview usage.
Instead, I will describe how Futhark models in \sasmodels are leveraged 
by \sasmodels' users.
% mere

@@

\section{Method and performance results}
% description of built-in sasmodels comparison functionality
\sasmodels contains a comparison module called \texttt{compare.py}, which
lets the user run different calculations on the same dataset.
This module has been the basis of all my performance tests, and indeed also
the module that has let me design Futhark models in the first place.

\subsection{Development method}
The development of Futhark for \sasmodels has been largely test-driven.
For any given already existing model in the \sasmodels catalogue that I have
wished to implement in Futhark, I have started out by defining an equivalent
model written in Futhark, and then used the comparison module to compare the
 runs of the original model versus the Futhark implementation.

@@

\subsection{Futhark vs. Python performance}
I have tested Futhark vs. Python performance on two different models; 
the \texttt{line} model and the \texttt{broad\_peak} model.

@@

\begin{figure}
    \begin{tikzpicture}

      \begin{axis}[
        mbarplot,
        title={1D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=3.5cm,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100, 0.24 ) (500, 0.24 ) (2500, 0.26 ) (5000, 0.26 ) (10000, 0.28 )};
      \addplot plot coordinates {(100, 0.93 ) (500, 0.94 ) (2500, 0.97 ) (5000, 0.95 ) (10000, 0.97 )};
      \legend{Current, Futhark}
      \end{axis}
    \end{tikzpicture}
    \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={2D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=4cm,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100, 0.31 ) (500, 2.78 ) (2500, 127.01 ) (5000, 501.31 ) (10000, 1970.14 )};
      \addplot plot coordinates {(100, 0.98 ) (500, 1.91 ) (2500, 22.73 ) (5000, 86.03 ) (10000, 338.10 )};

@@

\begin{figure}
      \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={1D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=3.5cm,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100,0.27) (500,0.30) (2500,0.45) (5000,0.63) (10000,1.22)};
      \addplot plot coordinates {(100,1.22) (500,1.23) (2500,1.23) (5000,1.25) (10000,1.27)};

      \legend{Current, Futhark}
      \end{axis}
    \end{tikzpicture}
    \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={2D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=4cm,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100,1.06 ) (500,21.50 ) (2500,675.12) (5000,2686.44 ) (10000,10687)};
      \addplot plot coordinates {(100,1.27 ) (500,2.27 ) (2500,27.48) (5000,103.66 ) (10000,409)};

@@

We see in the figure, that Futhark at 10000^2 datapoints now greatly outperforms
the Python model.
\subsubsection{Where does the performance gains come from?}
These comparisons are between Python/Numpy- and Futhark models.
Although Numpy can execute both of the models as vectorized calculations, 
it is still vastly outperformed by Futhark, which can run hundreds of threads 
simultaniously on the GPU.

## REFERENCER HER ##

\subsection{Futhark vs. OpenCL performance}
\subsubsection{The \texttt{DAB} model}
The DAB model is relatively simple, as it is described as $I(q) = \frac{L^3}{1 + (qL)^2}$,
requiring us to work with just one single variable.
  \begin{figure}
    \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={1D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=3.5cm,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100, 0.54 ) (500, 0.74 ) (2500, 0.76 ) (5000, 0.74 ) (10000, 0.77 )};
      \addplot plot coordinates {(100, 0.86 ) (500, 0.85 ) (2500, 0.87 ) (5000, 0.87 ) (10000, 0.90 )};

@@

      \end{axis}
    \end{tikzpicture}
    \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={2D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=4cm,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100, 0.78 ) (500, 1.61 ) (2500, 50.36 ) (5000, 196.19 ) (10000, 771.43 )};
      \addplot plot coordinates {(100, 0.90 ) (500, 1.75 ) (2500, 26.25 ) (5000, 101.59 ) (10000, 402.69 )};

@@

\subsubsection{The \texttt{Core Shell Parallelepiped} model}
The Core Shell Parallelepiped model is by far one of the most complex models
in \sasmodels' catalogue.
With it's more than ten parameters and it's two nested loops, it is 
a candidate model for the Futhark compiler to attempt to optimize.
  \begin{figure}
    \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={1D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=3.5cm,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100, 20.88 ) (500, 21.19 ) (2500, 21.99) (5000, 22.44 ) (10000, 42.99)};
      \addplot plot coordinates {(100, 2.68 ) (500, 3.67 ) (2500, 8.00 ) (5000, 13.18 ) (10000,23.90 )};

@@

      \end{axis}
    \end{tikzpicture}
    \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={2D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=4cm,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100, 0.83 ) (500, 2.17 ) (2500, 61.66 ) (5000, 232.32 ) (10000, 986.73 )};
      \addplot plot coordinates {(100, 2.04 ) (500, 2.88 ) (2500, 23.83 ) (5000, 91.87 ) (10000, 357.89 )};

@@

\subsubsection{Where does the performance gains come from?}
While it was expected to see Futhark outperform the strictly CPU based 
computations performed by Numpy in the Python models, it was more uncertain
how much (or even whether) Futhark would outperform OpenCL.

@@

% show core_shell-whatever; and elaborate on loop flattening into maps
%%% MASSER AF REFERENCER TIL TROELS' PAPERS

\subsection{Correctness of Futhark results}

\section{Discussion, future work}
% is this immediately useful? not on single fire runs with low nq
% can be used for data streams / bundling
% experience gains can lay groundwork for parallelising finding best match

% user directed: ease of use,
% (evt.) skønhedsfejl

\section{Conclusion}
% nævn at projektet er blevet præsenteret for en faktisk brugergruppe,
% og fik god og konstruktiv feedback
% det inspirerede til yderligere projekter/forbedringer indenfor sasmodels

>>>>

\documentclass[11pt]{article}
\usepackage[a4paper, hmargin={2.8cm, 2.8cm}, vmargin={2.5cm, 2.5cm}]{geometry}
\usepackage{eso-pic} % \AddToShipoutPicture
\usepackage{graphicx} % \includegraphics
\usepackage{listings} % \includegraphics
\usepackage{pgfplots}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{mdframed}
\usepackage{float}
\floatstyle{boxed}
\restylefloat{figure}

\lstset{
  basicstyle=\tiny,
  showspaces=false,
  showstringspaces=false,
  columns=flexible,
  breaklines=true
}

@@

\newcommand{\sasmodels}{\textit{SasModels}}
\newcommand{\sasview}{\textit{SasView}}
\newcommand{\iq}{\textit{I(q)}}
\newcommand{\foriq}{\texttt{for\_iq}}
\newcommand{\hasiq}{\texttt{has\_iq}}
\newcommand{\futhark}{\texttt{Futhark for SasModels}}
\newcommand{\barheight}{5cm}

\title{
  \vspace{3cm}
  \Huge{\futhark}\\
  \Large{Enhancing performance by rewriting kernels from Python/OpenCL to
    Futhark}
}

\author{
  \Large{Mikkel Storgaard Knudsen}\\
  \texttt{mikkelstorgaard@gmail.com}\\
  \\
  \Large{Supervisor}\\
  Martin Elsman\\
  \texttt{mael@di.ku.dk}
}

@@

\begin{document}

@@

\section{Introduction}
\subsection{Goal and motivation}
\sasview is a Small Angle Scattering (SAS) analysis package for the
analysis of 1D and 2D scattering data directly in inverse space.
As \sasview is given data, it performs its analyses by running the
input data through theory models provided by the standalone module
\sasmodels.

The \sasmodels models utilizes a set of kernels, which are implemented
in both Python and OpenCL. However, benchmark
comparisons\footnote{https://futhark-lang.org/performance.html}
between the parallel algorithm library \textit{Thrust}
(implemented in C++/CUDA), and the array programming language Futhark, shows
that there might severe speed improvements to gain by running these model
calculations in Futhark, instead of in OpenCL as they are now.

The goal of this project is to explore the potential speed gains obtained by
rewriting, and testing the resulting speed improvements of, a subset of the
\sasmodels models.
The subset will be chosen so that there are at least
one ``simple'' model, one model of intermediate complexity, and one complex
model in it.

If rewriting the models results in a significant\footnote{I.e. a speed increase
  in order of at least a magnitude}, speedup, it will reasonable to believe that 
  the rest of the models in \sasmodels can be rewritten for a similar
  speedup.

\section{Demonstration of use of finished project}
As \sasmodels is rarely used on its own, but rather as the backbone of 
\sasview's calculation backend, I am not going to write much about 
\sasview usage.
Instead, I will describe how Futhark models in \sasmodels are leveraged 
by \sasmodels' users.
måske ikke så meget mere her
% mere

@@

\section{Method}

% description of built-in sasmodels comparison functionality
\sasmodels contains a comparison module called \texttt{compare.py}, which
lets the user run different calculations on the same dataset.
This module has been the basis of all my performance tests, and indeed also
the module that has let me design Futhark models in the first place.

The development of Futhark for \sasmodels has been largely test-driven.
For any given already existing model in the \sasmodels catalogue that I have
wished to implement in Futhark, I have started out by defining an equivalent
model written in Futhark, and then used the comparison module to compare the
 runs of the original model versus the Futhark implementation.

@@

\section{Performance results}
\label{performance}
\subsection{Futhark vs. Python performance}
I have tested Futhark vs. Python performance on two different models; 
the \texttt{line} model and the \texttt{broad\_peak} model.

@@

\begin{figure}
    \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={1D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=\barheight,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100, 0.24 ) (500, 0.24 ) (2500, 0.26 ) (5000, 0.26 ) (10000, 0.28 )};
      \addplot plot coordinates {(100, 0.93 ) (500, 0.94 ) (2500, 0.97 ) (5000, 0.95 ) (10000, 0.97 )};
      \legend{Current, Futhark}
      \end{axis}
    \end{tikzpicture}
    \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={2D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=\barheight,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100, 0.31 ) (500, 2.78 ) (2500, 127.01 ) (5000, 501.31 ) (10000, 1970.14 )};
      \addplot plot coordinates {(100, 0.98 ) (500, 1.91 ) (2500, 22.73 ) (5000, 86.03 ) (10000, 338.10 )};

@@

\begin{figure}
      \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={1D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=\barheight,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100,0.27) (500,0.30) (2500,0.45) (5000,0.63) (10000,1.22)};
      \addplot plot coordinates {(100,1.22) (500,1.23) (2500,1.23) (5000,1.25) (10000,1.27)};

      \legend{Current, Futhark}
      \end{axis}
    \end{tikzpicture}
    \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={2D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=\barheight,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100,1.06 ) (500,21.50 ) (2500,675.12) (5000,2686.44 ) (10000,10687)};
      \addplot plot coordinates {(100,1.27 ) (500,2.27 ) (2500,27.48) (5000,103.66 ) (10000,409)};

@@

We see in the figure, that Futhark at $10000^2$ datapoints now greatly outperforms
the Python model.

\begin{mdframed}[
  frametitle={Why does Futhark perform faster than Python?},
  nobreak=true
  ]
These comparisons are between Python/Numpy- and Futhark models.
Although Numpy can execute both of the models as vectorized calculations, 
it is still vastly outperformed by Futhark, which can run hundreds of threads 
simultaniously on the GPU.

## REFERENCER HER ##
\end{mdframed}

\subsection{Futhark vs. OpenCL performance}
\subsubsection{The \texttt{DAB} model}
The DAB model is relatively simple, as it is described as $I(q) = \frac{L^3}{1 + (qL)^2}$,
requiring us to work with just one single variable.
  \begin{figure}
    \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={1D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=\barheight,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100, 0.54 ) (500, 0.74 ) (2500, 0.76 ) (5000, 0.74 ) (10000, 0.77 )};
      \addplot plot coordinates {(100, 0.86 ) (500, 0.85 ) (2500, 0.87 ) (5000, 0.87 ) (10000, 0.90 )};

@@

      \end{axis}
    \end{tikzpicture}
    \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={2D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=\barheight,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100, 0.78 ) (500, 1.61 ) (2500, 50.36 ) (5000, 196.19 ) (10000, 771.43 )};
      \addplot plot coordinates {(100, 0.90 ) (500, 1.75 ) (2500, 26.25 ) (5000, 101.59 ) (10000, 402.69 )};

@@

\subsubsection{The \texttt{Core Shell Parallelepiped} model}
The Core Shell Parallelepiped model is by far one of the most complex models
in \sasmodels' catalogue.
With it's more than ten parameters and it's two nested loops, it is 
a candidate model for the Futhark compiler to attempt to optimize.
  \begin{figure}
    \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={1D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=\barheight,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100, 20.88 ) (500, 21.19 ) (2500, 21.99) (5000, 22.44 ) (10000, 42.99)};
      \addplot plot coordinates {(100, 2.68 ) (500, 3.67 ) (2500, 8.00 ) (5000, 13.18 ) (10000,23.90 )};

@@

      \end{axis}
    \end{tikzpicture}
    \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={2D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=\barheight,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100, 0.83 ) (500, 2.17 ) (2500, 61.66 ) (5000, 232.32 ) (10000, 986.73 )};
      \addplot plot coordinates {(100, 2.04 ) (500, 2.88 ) (2500, 23.83 ) (5000, 91.87 ) (10000, 357.89 )};

@@

  \begin{mdframed}[
    frametitle={Why does Futhark perform faster than OpenCL?},
    nobreak=true]
While it was expected to see Futhark outperform the strictly CPU based 
computations performed by Numpy in the Python models, it was more uncertain
how much (or even whether) Futhark would outperform OpenCL.

@@

\# show core-shell-whatever; and elaborate on loop flattening into maps
%%% MASSER AF REFERENCER TIL TROELS' PAPERS
\end{mdframed}

\section{Discussion, future work}
\label{sec:discussion}
The performance tests in \ref{performance} unequivocally shows, that
there are speedups to gain by running Futhark kernels in favor of Python or
OpenCL kernels.
In general, Futhark starts surpassing the performance of competing kernels, when
sample sizes rise above n=INDSÆT N. However, these gains are currently
irrelevant for the most \sasview use cases, as almost all
experiments are carried out on sample sizes of between 100-500.
However, meetings with external project owner Wojciech Potrzebowski lead to the
suggestion that simple bundling of \sasview experiments would enable us
to leverage the Futhark speed improvements directly.

In example, one could imagine running an experiment that continuously sends
sensor data for processing in \sasview / \sasmodels.
Instead of running an experiment for each sample\footnote{of size n=100-500}
from the stream, one could let \sasmodels bundle these input samples in packages
of 10-50 samples
instead, and calculate these collectively to leverage the speedups that are
apparent on Futhark calculations on samples of larger sizes.

Please note, that sample sizes might have an upper bound, depending on the
memory available on the workstations running the experiments.

\subsection{Finding best match in \sasview}
\label{sec:best-match}
In addition to bundling input data, Wojciech also did suggest that the Futhark
speedups could be used to implement faster best-match-finding for \sasview
experiments.

Among other things, \sasview is used to test multiple models against a data set,
to find out which of the models that actually fits the data most closely.
Currently this is not done in an efficient manner.
Therefore, it would be interesting to implement a way of running multiple models
on the same data set in parallel.
And given the current Futhark vs. OpenCL results, the implementation should
pursued in Futhark, as it has already been shown to be optimal
(compared to OpenCL).

\subsection{Usability of Futhark models}
It is relevant to discuss how useful Futhark models are to the \sasview user
groups. For the users who merely wants to use the models, it is not necessary
to understand the implementation itself at all, as the user facing part of
the model is just the model info file, which is just the same as the model infos
for Python and OpenCL models \ref{sec:using-futhark}.

However, for the model developer, it is now also necessary to have at least a
rudimentary understanding of functional programming.
In example, developers will have to learn how to use maps in
places where they would usually use loops.

Furthermore, the current solution requires severe amounts of boilerplate code to
be included in the futhark model file.
This boilerplate is caused by the usage of Futhark's module system, and
therefore not strictly part of defining i.e. the \iq function of a model.
The boilerplate might become a distraction for the model developer, so
there might be a smaller educational challenge to solve here as well.

\section{Conclusion}
To answer the questions raised in this project definition, it was first
necessary to implement a \sasmodels kernel for Futhark models. This was done
successfully and in a very non-destructive way in the already existing
\sasmodels code base.

The kernel itself (\texttt{kernelfut.py}) depends on the existing code base, but
the total change of the original code is two lines: an extra case in the
switch statement that decides which kernel type to run for a given model.

For Futhark model performances increases themselves, the results in
\ref{performances}, tells us that there are performance increase of between
200- and 2000 percent, depending on the original model's language and
complexity.

These gains are only realised when the sample size for a given run is high
enough, but section \ref{sec:discussion} has proposals for \sasmodels extensions
that could help \sasmodels users to reach these sample sizes, by bundling
experiments.

\begin{mdframed}[
  frametitle={Presentation at \sasview{} Code Camp},
  nobreak=true
  ]
Together with my supervisor and Futhark designer Troels Henriksen,
I was invited for an hour-long presentation and discussion of \futhark{} at the
yearly \textit{SasView Code Camp}, held at the Copenhagen Bio Science Park.\\\\
There was about 20 \sasview{} users and developers attending at the presentation,
which was about the definition and implementation of \futhark{}, along with
a presentation of the result performance boosts for selected models.\\\\
The project was well-received by the audience, and there was an eager discussion
of whether and how it could be possible to really exploit the Futhark
optimizations.\\\\
The primary suggestion was to leverage Futhark to find best-fitting models,
as described in \ref{sec:best-match}, but bundling continuous input from
a sensor to exploit Futhark speed ups at high n was also a feature that many
of the attendants expressed an interest in.
\end{mdframed}

% nævn at projektet er blevet præsenteret for en faktisk brugergruppe,
% og fik god og konstruktiv feedback
% det inspirerede til yderligere projekter/forbedringer indenfor sasmodels
